partial: An R Package for Creating Partial Dependence Plots

### 2016-08-11 manuscript/greenwell.tex
minor edit
fixed typo

Let $\boldsymbol{x} = \left\{x_1, x_2, \dots, x_n\right\}$ represent the predictors in a model whose prediction function is $\widehat{f}\left(\boldsymbol{x}\right)$. If we partition $\boldsymbol{x}$ into an interest set, $\boldsymbol{z}_s$, and its compliment, $\boldsymbol{z}_{c} = \boldsymbol{x} \setminus \boldsymbol{z}_s$, then the partial dependence of the response on $\boldsymbol{z}_s$ is defined as
\begin{equation}
\label{eqn:pdf}
\bar{f}_s\left(\boldsymbol{z}_s\right) = \frac{1}{n}\sum_{i = 1}^n\widehat{f}\left(\boldsymbol{z}_s,\boldsymbol{z}_{i, c}\right),
\end{equation}
where $\boldsymbol{z}_{i, c}$ $\left(i = 1, 2, \dots, n\right)$ are the values of $\boldsymbol{z}_c$ that occur in the training sample; that is, weaverage out the effects of all the other predictors in the model. Let $x_1$ be the predictor variable of interest with unique values $\left\{x_{11}, x_{12}, \dots, x_{1k}\right\}$. The partial dependence of the response on $x_1$ can be constructed as follows:
\begin{enumerate}
  \item For $i \in \left\{1, 2, \dots, k\right\}$:
  \begin{enumerate}
    \item Create a copy of the training data and replace the original values of $x_1$ with the constant $x_{1i}$.
    \item Compute the vector of predicted values from the modified copy of the training data.
    \item Compute the average prediction to obtain $\bar{f}_1\left(x_{1i}\right)$.
  \end{enumerate}
  \item Plot the pairs $\left[x_{1i}, \bar{f}_1\left(x_{1i}\right)\right]$ for $i = 1, 2, \dotsc, k$.
\end{enumerate}
This can be quite computationally intensive since the algorithm involves $k$ passes over the training records. Fortunately, this can be parallelized quite easily. This algorithm can also be easily extended to larger subsets of 2 or more features as well.

>>>>

Let $\boldsymbol{x} = \left\{x_1, x_2, \dots, x_n\right\}$ represent the predictors in a model whose prediction function is $\widehat{f}\left(\boldsymbol{x}\right)$. If we partition $\boldsymbol{x}$ into an interest set, $\boldsymbol{z}_s$, and its compliment, $\boldsymbol{z}_{c} = \boldsymbol{x} \setminus \boldsymbol{z}_s$, then the partial dependence of the response on $\boldsymbol{z}_s$ is defined as
\begin{equation}
\label{eqn:pdf}
\bar{f}_s\left(\boldsymbol{z}_s\right) = \frac{1}{n}\sum_{i = 1}^n\widehat{f}\left(\boldsymbol{z}_s,\boldsymbol{z}_{i, c}\right),
\end{equation}
where $\boldsymbol{z}_{i, c}$ $\left(i = 1, 2, \dots, n\right)$ are the values of $\boldsymbol{z}_c$ that occur in the training sample; that is, we average out the effects of all the other predictors in the model. Let $x_1$ be the predictor variable of interest with unique values $\left\{x_{11}, x_{12}, \dots, x_{1k}\right\}$. The partial dependence of the response on $x_1$ can be constructed as follows:
\begin{enumerate}
  \item For $i \in \left\{1, 2, \dots, k\right\}$:
  \begin{enumerate}
    \item Create a copy of the training data and replace the original values of $x_1$ with the constant $x_{1i}$.
    \item Compute the vector of predicted values from the modified copy of the training data.
    \item Compute the average prediction to obtain $\bar{f}_1\left(x_{1i}\right)$.
  \end{enumerate}
  \item Plot the pairs $\left[x_{1i}, \bar{f}_1\left(x_{1i}\right)\right]$ for $i = 1, 2, \dotsc, k$.
\end{enumerate}
This can be quite computationally intensive since the algorithm involves $k$ passes over the training records. Fortunately, this can be parallelized quite easily. This algorithm can also be easily extended to larger subsets of 2 or more features as well.
